---
title: "Statistical Inference Project 1"
author: "fourlin88"
date: "Monday, June 15, 2015"
output: 
  html_document:
    keep_md: true
---

#Simulation to compare exponential distibution and Central Limit Theorem

## Requirement
Here is the requirement for this project. 
I need to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. It is required to set lambda = 0.2 for all of the simulations. I will investigate the distribution of averages of 40 exponentials. As another requirement, I will need to do a thousand simulations.

## Simulations

```{r}
# load neccesary libraries
library(ggplot2)

# set constants
lambda <- 0.2 # lambda for rexp
n <- 40 # number of exponetials
nSimulation <- 1000 # number of tests

# set the seed to create reproducability
set.seed(1125556665)

# run the test resulting in n x nSimulation matrix
expDist <- matrix(data=rexp(n * nSimulation, lambda), nrow=nSimulation)
expDistMeans <- data.frame(means=apply(expDist, 1, mean))
```

```{r echo=FALSE}
# plot the means
ggplot(data = expDistMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1) +   
  scale_x_continuous(breaks=round(seq(min(expDistMeans$means), max(expDistMeans$means), by=1)))
```

## Sample Mean versus Theoretical Mean

The expected mean $\mu$ of a exponential distribution of rate $\lambda$ is 

$\mu= \frac{1}{\lambda}$ 

```{r}
mu <- 1/lambda
print(mu)
```

Let $\bar X$ be the average sample mean of 1000 simulations of 40 randomly sampled exponential distributions.

```{r}
meanOfMeans <- mean(expDistMeans$means)
print(meanOfMeans)
```

The expected mean and the avarage sample mean are very close. 



## Sample Variance versus Theoretical Variance

The expected standard deviation $\sigma$ of a exponential distribution of rate $\lambda$ is 

$\sigma = \frac{1/\lambda}{\sqrt{n}}$ 


```{r}
sd <- 1/lambda/sqrt(n)
print(sd)
```

The variance $Var$ of standard deviation $\sigma$ is

$Var = \sigma^2$ 

```{r}
Var <- sd^2
print(Var)
```

suppose $Var_x$ be the variance of the average sample mean of 1000 simulations of 40 randomly sampled exponential distribution, and $\sigma_x$ the corresponding standard deviation.
```{r}
sd_x <- sd(expDistMeans$means)
print(sd_x)

Var_x <- var(expDistMeans$means)
print(Var_x)
```

From the results, standard deviations are very close. 

## Distribution

Let us now compare the population means & standard deviation with a normal distribution of the expected values. Please see the following graphs. 
Special lines were added in the graph to differentiate between the population means and the simulation means. 


```{r echo=TRUE}
# plot the means
ggplot(data = expDistMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1, aes(y=..density..), alpha=0.2) + 
  stat_function(fun = dnorm, arg = list(mean = mu , sd = sd), colour = "magenta", size=1) + 
  geom_vline(xintercept = mu, size=1, colour="magenta") + 
  geom_density(colour="blue", size=1) +
  geom_vline(xintercept = meanOfMeans, size=1, colour="blue") + 
  scale_x_continuous(breaks=seq(mu-3,mu+3,1), limits=c(mu-3,mu+3)) 
```

From the graph, the calculated distribution of means of random sampled exponantial distributions, overlaps quite nice with the normal distribution with the expected values based on the given lamba.


###References
Lecture notes, especially lecture number 7. 
